### 2023-02-20 paper/paper.rmd
Add `maxent` example

---
title: 'rang: Reconstructing reproducible R computational environments with ease'
tags:
  - R
  - reproducibility
  - docker
authors:
  - name: "Chung-hong Chan"
    orcid: 0000-0002-6232-7530
    affiliation: GESIS Leibniz-Institut für Sozialwissenschaften
  - name: "David Schoch"
    orcid: 0000-0003-2952-4812
    affiliation: GESIS Leibniz-Institut für Sozialwissenschaften    
abstract: |
  A complete description of the computational environment is often missing when researchers share their materials. Without such description, software obsolescence and missing system components can jeopardize computational reproducibility in the future, even when data and computer code are available. The R package rang is a complete solution for generating instructions for other researchers to completely reconstruct the computational environment at a specific time point. The reconstruction process, based on Docker, has been tested for R code as old as 2005 and does not depend on the long term availability of any commericial service. The description generated by rang satisfies the definition of a reproducible research compendium and can be shared as such.
keywords:
  - R
  - reproducibility
  - docker
date: 02 February 2023
year: 2023
bibliography: paper.bib
biblio-style: unsrt
output: rticles::arxiv_article
---

# Statement of need

The ability to reconstruct the computational environment is one of the most important aspects of reproducibility. In the realm of R, the important aspects are what and which version of installed R packages as well as the exact R version.

`renv` [@renvrpkg] (and its derivatives such as `jetpack` and its predecessor `packrat`) takes a similar approach to Python's `virtualenv` and Ruby's `Gem` to pin down the exact version of R packages using a "lock file". `containerit` [@nuest:2019] takes the current state of the computational environment and "containerize" it as a Dockerfile. But `containerit` does not pin down the exact version of R packages. Other solutions such as `groundhog` [@groundhogrpkg] and `checkpoint` [@checkpointrpkg] depend on the availability of The Microsoft R Application Network (MRAN), which will be shut down on July 1st, 2023.

These solutions are better for prospective usage, i.e. using them now to ensure the reproducibility of the current research for future researchers. `rang` mostly targets retrospective usage, i.e. using `rang` to reconstruct historical R computational environments which have not been completely declared. One can think of `rang` as an archaeological tool. In the following examples, `rang` is used to enable the reproducibility of published literature.

# Basic usage

There are two important functions of `rang`: `resolve()` and `dockerize()`.

`resolve()` queries various web services from the r-hub project of the R Consortium for information about R packages at a specific time point that is necessary for reconstructing a computational environment, e.g. recursive dependencies, system requirements, and R version. For instance, if there was a computational environment constructed on 2020-01-16 (called "snapshot date") with the several natural language processing R packages, `resolve()` can be used to resolve all the dependencies of these R packages. Currently, `rang` supports both CRAN and GitHub packages.

```r
library(rang)
graph <- resolve(pkgs = c("openNLP", "LDAvis", "topicmodels", "quanteda"),
                 snapshot_date = "2020-01-16")
graph
```

The resolved result is an S3 object called `rang`. The information contained in a `rang` can then be used to construct a computational environment in a similar manner as `containerit`, but with the packages and R versions pinned on the snapshot date. Then, `dockerize()` is used to generate the `Dockerfile` and other scripts in the `output_dir`. For R >= 3.1, the images from the Rocker project are used [@boettiger:2017:IR]. For R < 3.1 but >= 2.1, a custom image based on Debian Woody is used. As of writing, `rang` does not support R < 2.1, i.e. snapshot date earlier than 2005-04-19. The development to support R < 2.1 is in progress.

```r
dockerize(graph, output_dir = "docker")
```

There are two features of `dockerize()` that are important for future reproducibility.

1. By default, the container building process downloads source packages from CRAN and then compiles them from source. This step depends on the future availability of R packages on CRAN (which is extremely likely to be the case in the near future, given the continuous availability since 1997-04-23) [^1]. However, it is also possible to cache (or archive) the source packages now. The archived R packages can then be used instead during the building process.

[^1]: https://stat.ethz.ch/pipermail/r-announce/1997/000001.html

```r
dockerize(graph, output_dir = "docker", cache = TRUE)
```

2. It is also possible to install R packages in a separate library during the building process to isolate all these R packages from the main library.

```r
dockerize(graph, output_dir = "docker", cache = TRUE,
          lib = "anotherlibrary")
```

For the sake of completeness, the instructions for building and running the Docker container for Unix-like systems are included here.

```bash
cd docker
## might need to sudo
docker build -t rang .
docker run --rm --name "rangtest" -ti rang
```

# Case Studies

## quanteda JOSS paper

The software paper of the text analysis R package `quanteda` was published on 2018-10-06 [@benoit:2018]. In the paper, the following R code snippet is included.

```r
library("quanteda")
# construct the feature co-occurrence matrix
examplefcm <-
tokens(data_corpus_irishbudget2010, remove_punct = TRUE) %>%
tokens_tolower() %>%
tokens_remove(stopwords("english"), padding = FALSE) %>%
fcm(context = "window", window = 5, tri = FALSE)
# choose 30 most frequency features
topfeats <- names(topfeatures(examplefcm, 30))
# select the top 30 features only, plot the network
set.seed(100)
textplot_network(fcm_select(examplefcm, topfeats), min_freq = 0.8)
```

On 2023-02-08, this code snippet is not executable with the current version of `quanteda` (3.2.4). It is possible to install the "period appropriate" version of `quanteda` (1.3.4) using `remotes` on the current version of R (4.2.2). And indeed, the above code snippet can still be executed.

```r
remotes::install_version("quanteda", version = "1.3.4")
```

The issue is that installing `quanteda` 1.3.4 this way installs the latest dependencies from CRAN. `quanteda` 1.3.4 uses a deprecated (but not yet removed) function of `Matrix` (`as(<dgTMatrix>, "dgCMatrix")`). If this function were removed in the future, the above code snippet would not be executable anymore.

Using `rang`, one can query the version of `quanteda` on 2018-10-06 and create a Docker container with all the "period appropriate" dependencies. Here, the `rstudio` Rocker image is selected.

```r
library(rang)
graph <- resolve(pkgs = "quanteda",
                 snapshot_date = "2018-10-06",
                 os = "ubuntu-18.04")
dockerize(graph, output_dir = "quanteda_docker",
          image = "rstudio")
```

The above code snippet can be executed with the generated container without any problem.

```{r, fig.cap = "The code snippet running in a R 3.5.1 container created with rang", echo = FALSE}
knitr::include_graphics("quanteda_rstudio.png", dpi = 300)
```

## Psychological Science

 @cruewell:2023:WB evaluate the computational reproducibility of 14 articles published in *Psyhocological Science*. Among these articles, the paper by @hilgard:2019:NEG has been rated as having "package dependency issues".

All data and computer code are available from GitHub with the last commit on 2019-01-17 [^hilgard]. The R code contains a list of R packages used in the project as `library()` statements, including an R package on GitHub that is written by the main author of that paper. However, we identified one package (`compute.es`) that was not written in those `library()` statements but used with the namespace operator, i.e. `compute.es::tes()`. This undocumented package can be detected by `renv::dependencies()`.

[^hilgard]: https://github.com/Joe-Hilgard/vvg-2d4d

Based on the above information, one can run `resolve()` to obtain the dependency graph of all R packages on 2019-01-17.

```r
graph <- resolve(c("readxl", "rms", "ordinal", "tidyverse", "lubridate", "MBESS",
                   "censReg", "BayesFactor", "psych", "Joe-Hilgard/hilgard", "lsmeans",
                   "compute.es"),
                 snapshot_date = "2019-01-17")
```

When running `dockerize()`, one can take advantage of the `materials_dir` parameter to transfer the shared materials from @hilgard:2019:NEG into the Docker image.

```r
dockerize(graph, "hilgard", materials_dir = "vvg-2d4d", cache = TRUE)
```

We then built the Docker and launch a Docker container. For this container, we changed the entry point from R to bash so that the container goes to the Linux command shell instead.

```r
cd hilgard
docker build -t hilgard .
docker run --rm --name "hilgardcontainer" --entrypoint bash -ti hilgard
```

Inside the container, the materials are located in the `materials` directory. We used the following shell script to test the reproducibility of all R scripts.

```sh
cd materials
rfiles=(0_data_aggregation.R 1_data_cleaning.R 2_analysis.R 3_plotting.R)
for i in ${rfiles[@]}
do
	Rscript $i
	code=$?
	if [ $code != 0 ]
	then
		exit 1
	fi
done   
```

All R scripts ran fine inside the container and the figures generated are the same as the ones in @hilgard:2019:NEG.

## Political Analysis

The study by @trisovic:2022 evaluates the reproducibility of R scripts shared on Dataverse. They found that 75\% of R scripts cannot be successfully executed. Among these failed R scripts is an R script shared by @beck:2019:EGD.

This R script has been "rescued" by the author of the R package `groundhog` [@groundhogrpkg], as demonstrated in a blog post [^groundhog]. We were wondering if `rang` can also be used to rescue the concerned R script.

The date of the R script, as indicated on Dataverse, is 2018-12-12. This date is used as the snapshot date.

```r
graph <- resolve(c("foreign", "bife"), snapshot_date = "2018-12-12")
dockerize(graph, output_dir = "nat", materials_dir = "nathaniel")
```

```r
cd nat
docker build -t nat .
docker run --rm --name "natcontainer" --entrypoint bash -ti nat
```

Inside the container

```r
cd materials
Rscript fn_5.R
```

The same file can also be "rescued" by `rang`.

[^groundhog]: http://datacolada.org/100

## Recover a removed R package

The R package `maxent` introduces a machine learning algorithm with a small memory footprint and was available on CRAN until 2019. A software paper was published by the original authors in 2012 [@jurka:2012]. The R package was also used in some subsequent automated content analytic papers [e.g. @loercher:2017:D].

Despite the covert editing of the package by a staffer of CRAN [^evidence], the package was removed from CRAN in 2019 [^checkerror]. We attempted to install the second last (the original submitted version) and last (with covert editing) versions of `maxent` on R 4.2.2. Both of them didn't work.

Using `rang`, we are able to reconstruct a computational environment with R 2.15.0 (2012-03-30) to run all code snippets published in @jurka:2012 [^speed]. For removed CRAN packages, we strongly recommend querying the Github read-only mirror of CRAN instead (https://github.com/cran). It is because in this way, the resolved system requirements have a higher chance of being correct.

```r
maxent <- resolve("cran/maxent", "2012-06-10")
dockerize(maxent, "maxentdir", cache = TRUE)
```

[^checkerror]: https://cran-archive.r-project.org/web/checks/2019/2019-03-05_check_results_maxent.html

[^evidence]: https://github.com/cran/maxent/commit/9d46c6aad27a1f41a78907b170ddd9a586192be9

[^speed]: On an interesting historical note: The original paper reported, based on a benchmark, that "the algorithm is very fast; `maxent` uses only 135.4 megabytes of RAM and finishes in 53.3 seconds." On a modest computer in 2023 with a dockerized R 2.15.0, the benchmark finishes in 4 seconds.

# References