### 2023-06-09 paper/before-body.tex

$-- PLOS insert title and author within the main body
$-- Use this template to insert the necessary pieces before any content provided in .qmd file
\vspace*{0.2in}

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{$title$} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
\newline
\\
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
$by-author:_authors.tex()[, ]$
\\
\bigskip
$by-affiliation:_affiliations.tex()[\\ ]$
\bigskip

% Use the asterisk to denote corresponding authorship and provide email address in note below.
$for(by-author)$
$if(by-author.attributes.corresponding)$
* $by-author.email$
$endif$
$endfor$

\end{flushleft}

$if(abstract)$
\section*{Abstract}
$abstract$
$endif$

$if(author-summary)$
\section*{Author summary}
$author-summary$
$endif$

\linenumbers