### 2023-09-05 README.Rmd
Update handle [no ci]

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# rang <img src="man/figures/rang_logo.svg" align="right" height="200" />

<!-- badges: start -->
[![R-CMD-check](https://github.com/chainsawriot/rang/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/chainsawriot/rang/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The goal of rang (Reconstructing Ancient Number-crunching Gears) [^gesis] is to obtain the dependency graph of R packages at a specific time point.

Although this package can also be used to ensure the current R computational environment can be reconstructed by future researchers, this package gears towards reconstructing historical R computational environments which have not been completely declared. For the former purpose, packages such as [renv](https://github.com/rstudio/renv/), [groundhog](https://github.com/CredibilityLab/groundhog), [miniCRAN](https://github.com/andrie/miniCRAN), and [Require](https://github.com/PredictiveEcology/Require) should be used. One can think of rang as an archaeological tool.

To reconstruct a historical R computational environment, this package assumes only the availability of source packages online. The reconstruction procedures have been tested in several vintage versions of R.

Please cite this package as:

Chan CH, Schoch D (2023) rang: Reconstructing reproducible R computational environments. PLOS ONE [https://doi.org/10.1371/journal.pone.0286761](https://doi.org/10.1371/journal.pone.0286761)

## Installation

You can install the development version of rang like so:

``` r
remotes::install_github("gesistsa/rang")
```

Or the stable CRAN version

```r
install.packages("rang")
```

## Example

To obtain the dependency graph of R packages, use `resolve`. Currently, this package supports CRAN, Bioconductor, GitHub, and local packages.

```r
library(rang)
x <- resolve(pkgs = c("sna", "schochastics/rtoot", "S4Vectors"), snapshot_date = "2022-11-30")
```

```r
graph <- resolve(pkgs = c("openNLP", "LDAvis", "topicmodels", "quanteda"),
                 snapshot_date = "2020-01-16")
```

```{r, include = FALSE}
devtools::load_all()
graph <- readRDS("tests/testdata/graph.RDS")
```

```{r example1}
graph
```

```{r example2}
graph$sysreqs
```

```{r example 3}
graph$r_version
```

The resolved result is an S3 object called `rang` and can be exported as an installation script. The installation script can be execute on a vanilla R installation.

```r
export_rang(graph, "rang.R")
```

However, the execution of the installation script often fails (now) due to missing system dependencies and incompatible R versions. Therefore, the approach outlined below should be used.

## Recreate the computational environment via Rocker

A `rang` object can be used to recreate the computational environment via [Rocker](https://github.com/rocker-org/rocker). Please note that the oldest R version one can get from Rocker is R 3.1.0.

```r
dockerize(graph, "~/rocker_test")
```

Now, you can build and run the Docker container.

```bash
cd ~/rocker_test
docker build -t rang .
docker run --rm --name "rangtest" -ti rang
```

Using the above example, `sessionInfo()` outputs the following. You have successfully gone back to the pre-pandemic time.

```
R version 3.6.2 (2019-12-12)
Platform: x86_64-pc-linux-gnu (64-bit)
Running under: Debian GNU/Linux 10 (buster)

Matrix products: default
BLAS/LAPACK: /usr/lib/x86_64-linux-gnu/libopenblasp-r0.3.5.so

locale:
 [1] LC_CTYPE=en_US.UTF-8          LC_NUMERIC=C                 
 [3] LC_TIME=en_US.UTF-8           LC_COLLATE=en_US.UTF-8       
 [5] LC_MONETARY=en_US.UTF-8       LC_MESSAGES=C                
 [7] LC_PAPER=en_US.UTF-8          LC_NAME=en_US.UTF-8          
 [9] LC_ADDRESS=en_US.UTF-8        LC_TELEPHONE=en_US.UTF-8     
[11] LC_MEASUREMENT=en_US.UTF-8    LC_IDENTIFICATION=en_US.UTF-8

attached base packages:
[1] stats     graphics  grDevices utils     datasets  methods   base

other attached packages:
[1] topicmodels_0.2-9 LDAvis_0.3.2      openNLP_0.2-7     quanteda_1.5.2

loaded via a namespace (and not attached):
 [1] NLP_0.2-0           Rcpp_1.0.3          pillar_1.4.3       
 [4] compiler_3.6.2      tools_3.6.2         stopwords_1.0      
 [7] lubridate_1.7.4     lifecycle_0.1.0     tibble_2.1.3       
[10] gtable_0.3.0        lattice_0.20-38     pkgconfig_2.0.3    
[13] rlang_0.4.2         Matrix_1.2-18       fastmatch_1.1-0    
[16] parallel_3.6.2      openNLPdata_1.5.3-4 rJava_0.9-11       
[19] xml2_1.2.2          stringr_1.4.0       stats4_3.6.2       
[22] grid_3.6.2          data.table_1.12.8   R6_2.4.1           
[25] ggplot2_3.2.1       spacyr_1.2          magrittr_1.5       
[28] scales_1.1.0        modeltools_0.2-22   colorspace_1.4-1   
[31] stringi_1.4.5       RcppParallel_4.4.4  lazyeval_0.2.2     
[34] munsell_0.5.0       tm_0.7-7            slam_0.1-47        
[37] crayon_1.3.4    
```

### Caching R packages

One can also cache (or archive) the R packages from CRAN and Github at the time `dockerize` is executed. The cached R packages will then transfer to the container. Please note that system requirements (i.e. `deb` packages) are not cached.

```r
dockerize(graph, "~/rocker_test", cache = TRUE)
```

### Using alternative Rocker images

One can also select other Rocker versioned images: `rstudio`, `tidyverse`, `verse`, `geospatial`.

```r
dockerize(graph, "~/rocker_test", image = "rstudio")
```

`tidyverse`, `verse`, and `geospatial` are similar to the default (`r-ver`). For `rstudio`, one needs to build and launch it with:

```bash
cd ~/rocker_test
docker build -t rang .
docker run -p 8787:8787 -e PASSWORD=abc123 --rm --name "rangtest" -ti rang
```

With any browser, go to: `local:8787`. The default username is `rstudio`, password is as specified.

### Using Apptainer/Singularity containers

A `rang` object can be used to recreate the computational environment via [Rocker](https://github.com/rocker-org/rocker). Instead of Docker you can also use [Apptainer/Singularity](https://apptainer.org/). Please note that the oldest R version one can get from Rocker is R 3.1.0.

```r
apptainerize(graph, "~/rocker_test")
# singularize(graph, "~/rocker_test") # same function, as so far Apptainer is identical to Singularity
```

Now, you can build and run the Apptainer/Singularity container.

For Apptainer installation:

```bash
cd ~/rocker_test
apptainer build container.sif container.def
apptainer run container.sif R
```

For Singularity installation:

```bash
cd ~/rocker_test
sudo singularity build container.sif container.def
singularity run container.sif R
```

Using the above example, `sessionInfo()` outputs the following. You have successfully gone back to the pre-pandemic.

```
R version 3.6.2 (2019-12-12) -- "Dark and Stormy Night"
Copyright (C) 2019 The R Foundation for Statistical Computing
Platform: x86_64-pc-linux-gnu (64-bit)

R is free software and comes with ABSOLUTELY NO WARRANTY.
You are welcome to redistribute it under certain conditions.
Type 'license()' or 'licence()' for distribution details.

R is a collaborative project with many contributors.
Type 'contributors()' for more information and
'citation()' on how to cite R or R packages in publications.

Type 'demo()' for some demos, 'help()' for on-line help, or
'help.start()' for an HTML browser interface to help.
Type 'q()' to quit R.

> sessionInfo()
R version 3.6.2 (2019-12-12)
Platform: x86_64-pc-linux-gnu (64-bit)
Running under: Debian GNU/Linux 10 (buster)

Matrix products: default
BLAS/LAPACK: /usr/lib/x86_64-linux-gnu/libopenblasp-r0.3.5.so

locale:
 [1] LC_CTYPE=en_US.UTF-8       LC_NUMERIC=C
 [3] LC_TIME=en_US.UTF-8        LC_COLLATE=en_US.UTF-8
 [5] LC_MONETARY=en_US.UTF-8    LC_MESSAGES=C
 [7] LC_PAPER=en_US.UTF-8       LC_NAME=C
 [9] LC_ADDRESS=C               LC_TELEPHONE=C
[11] LC_MEASUREMENT=en_US.UTF-8 LC_IDENTIFICATION=C

attached base packages:
[1] stats     graphics  grDevices utils     datasets  methods   base

loaded via a namespace (and not attached):
[1] compiler_3.6.2
```

`apptainerize()`/`singularize()` functions work exactly the same as `dockerize()`, except you cannot cache Linux distribution rootfs.

### Apptainer/Singularity with RStudio IDE

To run RStudio IDE in Apptainer/Singularity container, some writeable folders and a config file have to be created locally:

```bash
mkdir -p run var-lib-rstudio-server .rstudio
printf 'provider=sqlite\ndirectory=/var/lib/rstudio-server\n' > database.conf
```

After that, you can run the container (do not run as `root` user, otherwise you will not be able to login to RStudio IDE).

Start instance (on default RSTUDIO port 8787):

```bash
apptainer instance start \
    --bind run:/run,var-lib-rstudio-server:/var/lib/rstudio-server,database.conf:/etc/rstudio/database.conf,.rstudio:/home/rstudio/.rstudio/ \
    container.sif \
    rangtest
```

Now open a browser and go to localhost:8787.
The default username is your local username, default password is 'set_your_password' (if you are using container generated by rang).

List running instances:

```bash
apptainer instance list
```

Stop instance:

```bash
apptainer instance stop rangtest
```

Start instance with custom port (e.g. 8080) and password:

```bash
apptainer instance start \
    --env RPORT=8080
    --env PASSWORD='set_your_password' \
    --bind run:/run,var-lib-rstudio-server:/var/lib/rstudio-server,database.conf:/etc/rstudio/database.conf,.rstudio:/home/rstudio/.rstudio/ \
    container.sif \
    rangtest
```

Run container with custom `rserver` command line:

```bash
apptainer exec \
    --env PASSWORD='set_your_password' \
    --bind run:/run,var-lib-rstudio-server:/var/lib/rstudio-server,database.conf:/etc/rstudio/database.conf,.rstudio:/home/rstudio/.rstudio/ \
    container.sif \
    /usr/lib/rstudio-server/bin/rserver \
    --auth-none=0 --auth-pam-helper-path=pam-helper \
    --server-user=$(whoami) --www-port=8787
```

If you run the container using `apptainer exec` command, you will have to kill the `rserver` process manually or Cmd/Ctrl+C from the running container to stop the server.

## Recreate the computational environment for R < 3.1.0

`rang` can still be used to recreate computational environments for R < 3.1.0. The Dockerfile generated is based on Debian Lenny (5.0) and the requested version of R is compiled from source. As of writing, this method works for R < 3.1.0 but not R < 1.3.1. The `image` parameter is ignored in this case.

```r
rang_rio <- resolve("rio", snapshot_date = "2013-08-28") ## R 3.0.1
dockerize(rang_rio, output_dir = "~/old_renviron")
```

## Acknowledgment

The logo of rang is a remix of [this](https://commons.wikimedia.org/wiki/File:Flag_of_the_Canary_Islands.svg) public domain image. The two dogs should be *Presa Canario*, the native dog breed on the islands of Gran Canaria and Tenerife.

---

[^gesis]: It stands for "R Archiving Nerds at GESIS". The package was previously named `gran`, but we decided to rename it to `rang` because there is another package named [gRAN](https://CRAN.R-project.org/package=GRANBase).