### 2023-10-30 paper/gesis_2023/index.qmd
Add gesis slides

---
title: "rang"
subtitle: "(Re)constructing reproducible R computational environment"
author: Chung-hong Chan, David Schoch
institute: GESIS
date: "2023-11-02"
date-format: "YYYY-MM-DD"
footer: "rang"
embed-resources: true
format: fakegesis-revealjs
---

## Case study

Are these analysis steps performed on the same dataset **consistently produce the same answer**?

```r
require(dplyr)
data <- data.frame(id = c(1,1,2), outcome = c(TRUE, TRUE, FALSE))
distinct(data, id) %>%
    select(outcome)
```

## On my own machine: dplyr 1.1.2

```{r}
#| echo: true
#| error: true
require(dplyr)
data <- data.frame(id = c(1,1,2), outcome = c(TRUE, TRUE, FALSE))
distinct(data, id) %>%
    select(outcome)
```

## Also on my own machine

![](dplyr.png)

## Time machine

![](dplyr1.png)

::: {style="font-size: 0.3em"}
Source: [https://dplyr.tidyverse.org/news/index.html#dplyr-050](https://dplyr.tidyverse.org/news/index.html#dplyr-050)
:::

::: {.notes}
Before Tidyverse (2016-09-09). Trump and Brexit too.
:::

## Method 1

Code rewrite

```{.r code-line-numbers="|3"}
require(dplyr)
data <- data.frame(id = c(1,1,2), outcome = c(TRUE, TRUE, FALSE))
distinct(data, id, .keep_all = TRUE) %>%
    select(outcome)
```

::: {.notes}
You'll need to edit your code when there are new breaking changes
:::

## Method 2

Install the old version

```{.r code-line-numbers="|1"}
remotes::install_version("dplyr", "0.4.3")
require(dplyr)
data <- data.frame(id = c(1,1,2), outcome = c(TRUE, TRUE, FALSE))
distinct(data, id) %>%
    select(outcome)
```

::: {.notes}
It installs the latest recursive dependencies
:::

## (Non-)method 3

Using ~~MRAN~~ Posit Public Package Manager (P3M)

```{.r code-line-numbers="|1,2"}
p3m <- "https://packagemanager.posit.co/cran/2016-01-10"
options(repos = c(REPO_NAME = p3m))
install.packages("dplyr") # install 0.4.3?

require(dplyr)
data <- data.frame(id = c(1,1,2), outcome = c(TRUE, TRUE, FALSE))
distinct(data, id) %>%
    select(outcome)
```

::: {.notes}
It won't work because P3M can only go as far back as 2017 Oct 10.
It installs the "time-appropriate" recursive dependencies. But it mixes old packages with your current system library.
Of course, you can use install.packages("dplyr", lib = "xyz")
:::

## (Non-)method 4

Using `renv` 1.0.0

```{.r code-line-numbers="|4"}
require(renv)
renv::init() ## restart
install.packages("dplyr")
renv::checkout(date = "2016-01-10")
renv::snapshot()
require(dplyr)
data <- data.frame(id = c(1,1,2), outcome = c(TRUE, TRUE, FALSE))
distinct(data, id) %>%
    select(outcome)
```

::: {.notes}
It won't work also because P3M can only go as far back as 2017 Oct 10.
:::

## Our solution: rang

* Date-based pinning (like P3M, `renv`)
* Create container images (like Rocker)
* Support CRAN, Bioconductor, GitHub, & local packages
* Resolve system dependencies, e.g. `libxml2`
* Depend on r-hub, CRAN, Bioconductor, GitHub API; doesn't use P3M
* Interoperability with `renv` (lockfile, project scanning)
* Can go as far back as **2001-08-31** (or R 1.3.1)

::: {.notes}
We are talking about Tony Blair and Bush 43 (and that's two weeks before 911)
:::

## Usage: Resolve

Resolve the dependency graph

```r
library(rang)
graph <- resolve(pkgs = c("dplyr"), snapshot_date = "2016-01-10")
graph
```

```{r}
#| echo: false
library(rang)
readRDS("dplyr.RDS")
```

## Usage: Dockerize

```r
dockerize(rang = graph,
          output_dir = "~/dplyr_example",
          no_rocker = TRUE,  # recommended
          cache = TRUE) # recommended
```

## What's inside

```text
├── cache
│   ├── debian
│   │   └── rootfs.tar.xz
│   ├── rpkgs
│   │   ├── assertthat_0.1.tar.gz
│   │   ├── BH_1.60.0-1.tar.gz
│   │   ├── DBI_0.3.1.tar.gz
│   │   ├── dplyr_0.4.3.tar.gz
│   │   ├── lazyeval_0.1.10.tar.gz
│   │   ├── magrittr_1.5.tar.gz
│   │   ├── R6_2.1.1.tar.gz
│   │   └── Rcpp_0.12.2.tar.gz
│   └── rsrc
│       └── R-3.2.3.tar.gz
├── compile_r.sh
├── Dockerfile
├── rang.R
└── README

4 directories, 14 files
```

## Dockerfile

```Dockerfile
FROM scratch
ADD /cache/debian/rootfs.tar.xz /
ENV TZ UTC
RUN ln -snf /usr/share/zoneinfo/$TZ /etc/localtime \
	&& echo $TZ > /etc/timezone \
	&& apt-get update -qq \
	&& apt-get install wget locales build-essential r-base-dev  -y
ENV RANG_PATH /rang.R
ENV COMPILE_PATH /compile_r.sh
ENV CACHE_PATH /cache
COPY rang.R /rang.R
COPY compile_r.sh /compile_r.sh
COPY cache/rpkgs /cache/rpkgs
COPY cache/rsrc /cache/rsrc
RUN apt-get update -qq \
	&& apt-get install -y libpcre3-dev zlib1g-dev pkg-config libcurl4-openssl-dev
RUN bash $COMPILE_PATH 3.2.3
CMD ["R"]
```

## Build and launch

```sh
cd ~/dplyr_example
docker build -t dplyr .
docker run --rm --name "dplyr" -ti dplyr
```

. . .

![](dplyr.png)

# Summary

* code edit
* Install old version
* renv
* P3M
* rang
* New development: evercran (CRAN snapshot and versioned Docker images)